% ======================================================================================
%  Tao 2019 — "Almost all orbits of the Collatz map attain almost bounded values"
%  Lean 4 formalization blueprint (arXiv:1909.03562v5).
%
%  This is the reviewer-facing proof map: LaTeX narrative + `\lean{}` bindings (to the
%  canonical declaration names fixed in SKELETON-SPEC.md) + `\uses{}` dependency edges
%  (following the "depends" column of BLUEPRINT.md's node ledger) + `\leanok`/`\notready`
%  status.  Node ids are the labels: S1--S4 (support), C1--C10 (core spine), X1--X11
%  (the §7 crux).  The prose statement of each node is transcribed from BLUEPRINT.md's
%  descriptions and SKELETON-SPEC.md's Lean statements; where the paper equation is not
%  reproduced verbatim here a `% TODO-PRECISE` marks the coarse spot.
%
%  STATUS DISCIPLINE (updated 2026-07-12, judge pass 8):
%    * Statement-`\leanok` is set on nodes whose `\lean{}`-bound statements are landed,
%      compiled, and judge-ratified.  The nodes with no pinned Lean statement yet
%      (C8, X1, X5) keep `\notready`.  PROOF-`\leanok` (green background) is set ONLY on a
%      dated judge-run `#print axioms` verification of the bound decls -- the manual form
%      of the future BlueprintAudit reconciler; never on a worker's claim alone.
%      (Proof-verified so far: X3, S3, X6 -- see EXECUTABILITY.md judge log.)
%    * Nodes with a canonical Lean name in SKELETON-SPEC.md carry `\lean{...}`.  Nodes the
%      spec leaves un-named carry no `\lean{}` and are described in prose; they will be
%      bound once named.
%    * Every node whose WORK is open carries a `\lapsrisk{laps}{risk}{conf}` campaign
%      estimate mirroring BLUEPRINT.md §2's ledger (laps = expected treadmill laps,
%      risk word from the as-stated confidence: >=85 low, 75-80 medium, <=70 high).
%      Drop the line only when the node's work is done; re-sync numbers from the ledger.
%      NB the badge is per-node WORK, not the dep-graph fill: leanblueprint colors a
%      *definition* node green (dark green when its whole ancestor cone is green) as soon
%      as its bound defs compile, while the node's support lemmas may still be open -- so
%      "dark green + lap badge" on a definition node is intentional, not stale (e.g. S1's
%      Prob/Basic.lean sorries, C6's Statement.lean sorries).
% ======================================================================================

% Campaign estimate annotation (one per incomplete node). Defined here so both the
% plasTeX web build (web.tex) and the pdf build (print.tex) pick it up via \input.
\newcommand{\lapsrisk}[3]{\par\smallskip\noindent\textit{\textbf{Campaign estimate:}
#1 treadmill laps; risk \textbf{#2} (#3 confidence the node completes as stated).}}

\chapter{Program overview}

This is a blueprint for a \emph{full Lean~4 formalization of Terence Tao's 2019 theorem}
(\texttt{arXiv:1909.03562v5}) that almost all Collatz orbits attain almost bounded values.
The summit is \textbf{Theorem~1.3}: for every \(f:\mathbb{N}\to\mathbb{R}\) with
\(f(N)\to\infty\), one has \(\mathrm{Colmin}(N) < f(N)\) for almost all \(N\) in the sense of
\emph{logarithmic density}, where \(\mathrm{Colmin}(N)\) is the minimum of the Collatz orbit
of \(N\).  As of this writing the theorem is formalized in \emph{no} proof assistant, by any
route; a Lean proof is a first.

\section{The spine}

The whole argument is a chain of reductions.  Read right-to-left, each arrow is
``is proved from'':
\[
  \underbrace{\text{Thm 1.3}}_{\text{main}}
  \Longleftarrow \text{Thm 1.6}
  \Longleftarrow \underbrace{\text{Thm 3.1}}_{\text{quantitative}}
  \Longleftarrow \underbrace{\text{Prop 1.11}}_{\text{stabilization}}
  \Longleftarrow \big[\,\text{Prop 1.9} + \text{Prop 1.14}\,\big]
  \Longleftarrow \underbrace{\text{Prop 1.17}}_{\text{character decay}}
  \Longleftarrow \text{Prop 7.1}
  \Longleftarrow \text{Prop 7.3}
  \Longleftarrow \text{Prop 7.8}.
\]
The passage from the analytic main theorem to a purely arithmetic statement about the
\emph{Syracuse} map (the odd-to-odd accelerated Collatz map) happens at
Thm~1.6/Thm~3.1 (node~\ref{C6}), and drives everything down to a single Fourier-analytic
input, the \textbf{character-sum decay} Prop~1.17 (node~\ref{X11}).  Prop~1.17 is proved in
§7 by a renewal-process argument played against the geometry of the ``black'' set of the
phase functions --- the deepest part of the paper.

\section{The three risk kernels}

Almost all of the completion risk concentrates in three clusters; everything outside them is
standard treadmill fare (counting, harmonic sums, \(\mathbb{Z}/3^n\mathbb{Z}\) arithmetic,
PMF calculus).
\begin{itemize}
  \item \textbf{S3 --- the local 2-D bound} (Lemma~2.2, node~\ref{S3}): Gaussian-type local
        limit upper bounds for sums of the lattice distributions.  Classical analysis;
        long but self-contained via explicit formulas and exponential tilting (D5).
  \item \textbf{X3 --- the triangle structure} (Lemma~7.4, node~\ref{X3}): the black set of
        the phase function is a disjoint union of well-separated triangles.  Finite,
        elementary, delicate case analysis.
  \item \textbf{X8/X10 --- renewal versus triangles} (Case~2 and Lemma~7.10,
        nodes~\ref{X8},~\ref{X10}): the paper's pinnacle.  The design decision D6 finitizes
        the infinite renewal process into an explicit downward recursion \(Q\), turning the
        stopping-time measure theory into strong inductions over that recursion.
\end{itemize}
The critical path is
\(\text{S3}\to\text{X6}\to\{\text{X8},\text{X10}\}\to\text{X11}\to\text{C10}\to\text{C9}
  \to\text{C6}\).

\section{Global modelling decisions}

The Lean development departs from the paper's presentation in eight deliberate ways
(full rationale in \texttt{BLUEPRINT.md} §0); the ones that shape the statements below are:
\begin{itemize}
  \item[\textbf{D1}] \emph{Probability is discrete.}  Every random variable is a \texttt{PMF}
        on a countable type; expectations are \texttt{tsum}s and total variation is
        \(d_{\mathrm{TV}}(p,q)=\sum_a |p(a)-q(a)|\).  No measure theory (node~\ref{S1}).
  \item[\textbf{D2}] \(\mathbb{Z}[1/2]\) \emph{is eliminated.}  The paper's offset map is
        replaced by its integer multiple \(\texttt{fnat}\,n\,a = \sum_{m<n} 3^{n-1-m}
        2^{\,a_{[1,m]}}\), so the iteration identity (1.7)\(\times 2^{|a|}\) lives entirely
        in \(\mathbb{N}\) (node~\ref{C2}).
  \item[\textbf{D3}] \emph{Asymptotics are reified.}  \(X\ll Y\) becomes
        \(\exists C>0,\ X\le C\,Y\); \(\ll_A\) becomes \(\forall A>0,\exists C>0\).  No
        \texttt{IsBigO}/filters in load-bearing statements (uniformity in \(n,\xi\) is
        exactly what filter-\(O\) would hide).  This constrains \emph{statements} only, and
        does not by itself make a constant effective: a proof may still \texttt{obtain} its
        witness from a rate-free limit and satisfy D3 while defeating its purpose.  Where a
        constant is meant to be effective the witness must be an explicit term; since
        \texttt{Exists} is a \texttt{Prop}, an inexplicit one cannot be recovered downstream.
        See \texttt{notes/effective-constants.md}.
  \item[\textbf{D4}] \emph{The small constant is a numeral.}  \(\varepsilon := 10^{-4}\)
        (candidate; validated per-site).
  \item[\textbf{D6}] \emph{The infinite renewal process is finitized} by the downward
        recursion \(Q\) (node~\ref{X4}) --- the single most important de-risking decision.
\end{itemize}
Throughout, tuples are \(0\)-indexed (\(a:\mathrm{Fin}\,n\to\mathbb{N}\); paper index \(i\)
is Lean index \(i-1\)), and \(a_{[1,m]}\) denotes the prefix sum \(\sum_{i<m} a_i\).


\chapter{The support layer}\label{chp:support}

Discrete-probability, generating-function, and Fourier scaffolding used everywhere below.

\section{PMF calculus}

\begin{definition}[Discrete-probability calculus]\label{S1}
  \lean{PMF.dTV, PMF.expect, PMF.cexpect, PMF.iid}
  \leanok
Total variation distance \(d_{\mathrm{TV}}(p,q) := \sum_a' |(p\,a).\mathrm{toReal} - (q\,a).\mathrm{toReal}|\),
real and complex expectations \(\mathbb{E}_p[f] := \sum_a' (p\,a).\mathrm{toReal}\cdot f(a)\),
and the \(n\)-fold i.i.d.\ product \(p^{\otimes n}\) on \(\mathrm{Fin}\,n\to\alpha\)
(node D1).  Together with the coupling inequality (paper (1.10)), stated in indicator-expectation
form to avoid \texttt{OuterMeasure} friction: for any event \(E\subseteq\alpha\),
\[
  \bigl|\mathbb{E}_p[\mathbf 1_E] - \mathbb{E}_q[\mathbf 1_E]\bigr| \le d_{\mathrm{TV}}(p,q).
\]
Basic algebra of \(d_{\mathrm{TV}}\) (symmetry, non-negativity) and the support/apply lemmas of
the i.i.d.\ product are part of this node.
\lapsrisk{8--15}{low}{90\%}
\end{definition}

\section{The geometric family and local bounds}

\begin{definition}[Geometric and Pascal laws; negative binomial]\label{S2}
  \lean{geomHalf, geomQuarter, pascal, pascalNe3}
  \uses{S1}
  \leanok
The lattice laws driving the Syracuse valuation: \(\mathrm{Geom}(2)\) with
\(\mathbb{P}(a)=2^{-a}\) for \(a\ge 1\) (\texttt{geomHalf}); \(\mathrm{Geom}(4)\)
(\texttt{geomQuarter}); \(\mathrm{Pascal}\), the law of \(a_1+a_2\) with
\(a_i\sim\mathrm{Geom}(2)\), i.e.\ \(\mathbb{P}(b)=(b-1)2^{-b}\) for \(b\ge 2\)
(\texttt{pascal}); and its conditioning off \(b=3\) (\texttt{pascalNe3}).  The load-bearing
computation is the exact \textbf{negative-binomial point mass} for the sum of \(n\) i.i.d.\
\(\mathrm{Geom}(2)\) variables,
\[
  \mathbb{P}\!\Bigl(\textstyle\sum_{i<n} a_i = L\Bigr)
    = \binom{L-1}{n-1}\,2^{-L}, \qquad L\ge n\ge 1
\]
(compositions of \(L\) into \(n\) positive parts), together with the moment generating
functions and the one-dimensional Chernoff upper-tail bounds derived from them.
\lapsrisk{6--12}{low}{90\%}
\end{definition}

\begin{lemma}[Local 2-D Gaussian-type bound; Lemma 2.2]\label{S3}
  \lean{Gweight, geomHalf_local_bound, geomQuarter_local_bound, pascal_local_bound, geomHalf_tail_bound, geomQuarter_tail_bound, pascal_tail_bound, hold_local_bound, hold_tail_bound}
  \uses{S2}
  \leanok
\textbf{Risk kernel 1.}  Lemma~2.2(i)(ii): local-limit / Gaussian-type \emph{upper} bounds on
the point masses of sums of \(k\) i.i.d.\ copies of the relevant lattice distributions
(\(\mathrm{Geom}(2)\), \(\mathrm{Geom}(4)\), \(\mathrm{Pascal}\), and the holding law
\(\mathrm{Hold}\) of §7), uniform in the shift, of the shape
\(\mathbb{P}(S_k = v)\ll k^{-d/2}\) with \(d\) the lattice dimension; together with the weight
function \(G_n\) (paper (2.2)).  Proved not by the paper's contour/Morera route but by explicit
generating functions plus real-variable exponential tilting (design decision D5): for the
\(d=2\) holding sums this is the circle-method form
\(\mathbb{P}(S_k=v)=(2\pi)^{-2}M(\lambda)^k e^{-\lambda\cdot v}\int_{[-\pi,\pi]^2}
|\varphi_\lambda(t)|^k\,dt\) with \(\varphi_\lambda,M\) explicit rational functions.
Statements judge-ratified 2026-07-12 vs pp.14--15: per-instance forms in
\texttt{Prob/LocalInstances.lean} (scalar) and \texttt{Sec7/HoldLocal.lean} (\(d=2\) Hold,
with the correct \((n+1)^{-1}\) prefactor and product norm fed to the scalar \(G\)); the
\(G_{1+n}\)-for-\(G_n\) index is constants-equivalent (in-paper precedent: Lemma 7.7).
\end{lemma}
\begin{proof}
  \leanok
Risk kernel 1 CLOSED (campaign laps 22--45; judge pass 5, 2026-07-12).  All eight
obligations --- (i) and (ii) for \(\mathrm{Geom}(2)\), \(\mathrm{Geom}(4)\),
\(\mathrm{Pascal}\), \(\mathrm{Hold}\) --- machine-checked via the D5 route (circle-method
inversion, character decay, exponential tilting, second-order MGF bounds with exact means,
\(\lambda\)-clip Chernoff assembly), plus the generic engines
\texttt{iidSum\_nat\_local\_of\_quad} / \texttt{iidSum\_nat\_tail\_of\_quad}.  Judge-run
\texttt{\#print axioms} on all ten declarations (2026-07-12):
[propext, Classical.choice, Quot.sound] --- nothing extra.
\end{proof}

\section{Fourier analysis on \(\mathbb{Z}/3^n\mathbb{Z}\)}

\begin{definition}[Additive characters and the oscillation functional]\label{S4}
  \lean{eC, osc}
  \leanok
The additive character \(e(\theta):=\exp(2\pi i\theta)\) on \(\mathbb{Q}\), the discrete
Fourier transform and Parseval identity on \(\mathbb{Z}/3^n\mathbb{Z}\), and the
\textbf{oscillation functional} \(\mathrm{Osc}_{m,n}\) (paper (1.24)) measuring how far a
weight \(c:\mathbb{Z}/3^n\mathbb{Z}\to\mathbb{R}\) is from being constant on the fibres of the
projection to \(\mathbb{Z}/3^m\mathbb{Z}\):
\[
  \mathrm{Osc}_{m,n}(c)=\sum_{Y}\Bigl|\,c(Y)-3^{\,m-n}\!\!\sum_{Y'\mapsto \bar Y} c(Y')\,\Bigr|,
  \qquad m\le n.
\]
Includes the Remark~1.18 triangle inequality that lets \(\mathrm{Osc}\) be controlled scale by
scale.
\lapsrisk{4--8}{low}{85\%}
\end{definition}


\chapter{The Collatz--Syracuse dictionary and logarithmic density}\label{chp:dictionary}

\begin{definition}[Collatz and Syracuse maps; the reduction (1.2)]\label{C1}
  \lean{col, colMin, oddPart, syr, syrMin, colMin_eq_syrMin_oddPart}
  \leanok
The Collatz map \(\mathrm{col}(N)=3N+1\) if \(N\) is odd and \(N/2\) otherwise; its orbit
minimum \(\mathrm{Colmin}(N)=\inf_k \mathrm{col}^{[k]}(N)\).  The odd part
\(\mathrm{oddPart}(N)=N/2^{v_2(N)}\), the \textbf{Syracuse map}
\(\mathrm{syr}(N)=\mathrm{oddPart}(3N+1)\) (odd \(\to\) odd), and its orbit minimum
\(\mathrm{Syrmin}\).  The elementary reduction (paper (1.2)) that lets the entire problem be
studied on the odd numbers:
\[
  0<N \;\Longrightarrow\; \mathrm{Colmin}(N) = \mathrm{Syrmin}(\mathrm{oddPart}\,N).
\]
Also the cheap odd-preservation and positivity facts (\(\mathrm{syr}\) maps odds to odds, etc.).
\lapsrisk{3--6}{low}{95\%}
\end{definition}

\begin{lemma}[Valuation vector, integer offset, and the iteration identity (1.7)]\label{C2}
  \lean{valVec, fnat, syr_iterate_key, valVec_unique, syr_iterate_odd}
  \uses{C1}
  \leanok
The \(n\)-step \textbf{valuation vector} (paper (1.8))
\(\vec a^{(n)}(N)_i = v_2\bigl(3\,\mathrm{syr}^{[i]}(N)+1\bigr)\), the integerified offset
(design D2) \(\texttt{fnat}\,n\,a = \sum_{m<n} 3^{\,n-1-m}\,2^{\,a_{[1,m]}}\), and the
load-bearing \textbf{iteration identity}, the paper's (1.7) multiplied through by \(2^{|a|}\)
so it holds in \(\mathbb{N}\): for odd \(N\), with \(a=\vec a^{(n)}(N)\),
\[
  2^{\,a_{[1,n]}}\cdot \mathrm{syr}^{[n]}(N) = 3^n N + \texttt{fnat}\,n\,a .
\]
(By induction on \(n\); the step is \(2^{a_{n+1}}\mathrm{syr}(\mathrm{syr}^{[n]}N)
= 3\,\mathrm{syr}^{[n]}N + 1\).)  Together with Lemma~2.1: the valuation vector is the
\emph{unique} tuple with all entries \(\ge 1\) making the affine quotient odd.
\textbf{COMPLETE (judge pass 21, 2026-07-13).}  \texttt{valVec\_unique} (Lemma 2.1 as a
hypothesis-guarded iff: divisibility \(\wedge\) odd quotient \(\leftrightarrow\)
\(a = \texttt{valVec}\), for tuples with entries \(\ge 1\)) was proved by an external
Codex session and judge-verified: statement \textbf{RATIFIED vs Lemma 2.1 p.14}
(RATIFY-2 resolved --- the iff matches the paper's uniqueness claim; the paper's
membership half, \(\vec a^{(n)} \in (\mathbb{N}+1)^n\) itself, is an unformalized
one-line companion, consumer-safe since candidate tuples carry the hypothesis by
construction); proof route is the paper's own last-entry induction; dated judge runs
2026-07-13 on \texttt{syr\_iterate\_key}, \texttt{valVec\_unique},
\texttt{syr\_iterate\_odd} all exactly
\([\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}]\).
\end{lemma}
\begin{proof}
  \leanok
  Forward: induction on \(n\) peeling the last entry (the paper's own route) --- the
  truncated vector satisfies the \(n\)-step condition (recursion identities for
  \texttt{pre}/\texttt{fnat} under truncation), the quotient is identified with
  \(\mathrm{syr}^{[n]}(N)\) via the iteration identity, and the last entry is extracted
  as \(\nu_2(3\,\mathrm{syr}^{[n]}(N)+1)\).  Backward: the iteration identity gives
  divisibility, and the quotient is \(\mathrm{syr}^{[n]}(N)\), odd since Syracuse
  iterates of odd numbers stay odd (\texttt{syr\_iterate\_odd}).
\end{proof}

\begin{definition}[Logarithmic density and the reduction Thm 1.6 \(\Rightarrow\) Thm 1.3]\label{C3}
  \lean{logProb, HasLogDensity, AlmostAllPos, AlmostAllOdd}
  \uses{C1}
  \leanok
The logarithmic-density apparatus: \texttt{logSum}, \texttt{logProb}, the sampling window
\texttt{Finset.Icc 1 x}, the predicate \texttt{HasLogDensity}, and the ``almost all''
quantifiers \texttt{AlmostAllPos} and its odd-window form \texttt{AlmostAllOdd}
(logarithmic density restricted to \(2\mathbb{N}+1\)).  With the harmonic-sum integral tests
(paper (5.25)(5.26)) and the elementary splitting reducing the even-index statement
Thm~1.6 to the main theorem Thm~1.3 (node~\ref{C6}) by passing to odd parts.
\lapsrisk{6--12}{low}{85\%}
\end{definition}


\chapter{The Syracuse random variable and its valuation}\label{chp:syracuse}

\begin{definition}[The Syracuse random variable \(\mathrm{Syrac}(\mathbb{Z}/3^n\mathbb{Z})\)]\label{C4}
  \lean{syracZ, syracZ_map_cast}
  \uses{C2, S1}
  \leanok
The law \texttt{syracZ}\,\(n\) on \(\mathbb{Z}/3^n\mathbb{Z}\) (paper (1.21)/(1.26), in the
\emph{reversed} variable order deliberately, per footnote~6): the pushforward of
\(\mathrm{Geom}(2)^{\otimes n}\) under
\[
  a \;\longmapsto\; \sum_{j<n} 3^{\,j}\,(2^{-1})^{\,a_{[1,j+1]}} \in \mathbb{Z}/3^n\mathbb{Z},
\]
where \(2^{-1}\) is the inverse of the unit \(2\) mod \(3^n\).  With the projection
compatibility (paper (1.22)): \texttt{syracZ}\,\(n\) pushes forward to \texttt{syracZ}\,\(k\)
under \(\mathbb{Z}/3^n\mathbb{Z}\to\mathbb{Z}/3^k\mathbb{Z}\) for \(k\le n\); and the recursion
Lemma~1.12 (Euler's theorem plus a geometric-series normalization
\((1-2^{-2\cdot 3^n})^{-1}\)).
\lapsrisk{5--10}{low}{85\%}
\end{definition}

\begin{proposition}[Valuation \(\approx\) geometric; Prop 1.9 and the §4 tail bound]\label{C5}
  \lean{valuation_dist, valuation_tail, unifOddMod, valVec_pos}
  \uses{C2, S2}
  \leanok
\textbf{Proposition 1.9.}  There are \(c_1,C>0\) such that for all \(n,n'\) and every law
\(X\) supported on odd numbers that is \(\ll 2^{-n'}\)-close in total variation to uniform
modulo \(2^{n'}\), provided \((2+c_0)n\le n'\),
\[
  d_{\mathrm{TV}}\!\bigl(X_*\vec a^{(n)},\ \mathrm{Geom}(2)^{\otimes n}\bigr)
    \le C\,2^{-c_1 n}.
\]
I.e.\ the Syracuse valuation vector of an equidistributed-mod-\(2^{n'}\) input is
exponentially close to \(n\) independent geometric variables.  Includes the §4 tail bound
Lemma~4.1 controlling \(\mathbb{P}(a_{[1,n]}\ge n')\).
\textbf{COMPLETE (judge pass 22, 2026-07-13).}  \texttt{valuation\_dist} (Prop 1.9) and
\texttt{valuation\_tail} (Lemma 4.1) proved by an external Codex session and
judge-verified: pinned statements character-untouched (constituent
\texttt{unifOddMod} body unchanged; statements re-read vs Prop 1.9 p.7 / Lemma 4.1
p.22 --- the \(\forall c_0,K\,\exists c_1,C\) quantifier structure matches the paper's
``constants may depend on \(c_0\)'' exactly).  The tail proof consumed the
long-parked (1.10) hole \texttt{PMF.abs\_expect\_indicator\_sub\_le\_dTV}
(Prob/Basic), caught by the dated axiom run and discharged by the judge same pass.
Dated runs 2026-07-13 on both headliners + \texttt{valVec\_pos},
\texttt{fnat\_mod\_two\_of\_pos}, \texttt{unifOddMod\_map\_valVec\_apply},
\texttt{truncated\_val\_dTV\_le}, \texttt{PMF.dTV\_map\_le} --- all exactly
\([\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}]\).
\end{proposition}
\begin{proof}
  \leanok
  Route (paper §4, with one sound reordering): the event
  \(\vec a^{(n)}(N) = \vec a\) is classified as a single residue of \(N\) modulo
  \(2^{a_{[1,n]}+1}\) (via Lemma 2.1, node~\ref{C2}); pushing the
  equidistribution hypothesis through the truncation \(2^{n'} \to 2^{|a|+1}\)
  (data processing, \texttt{PMF.dTV\_map\_le}) matches the truncated valuation
  law with \(\mathrm{Geom}(2)^{\otimes n}\) exactly below level \(n'\); the
  overflow \(|a| \ge n'\) is controlled on the geometric side by the S2 tail
  engine.  Lemma 4.1 is then \emph{derived} from Prop 1.9 plus the geometric
  tail (reverse of the paper's order --- sound and non-circular, since
  \(\mathbb{P}_P(E) \le \mathbb{P}_Q(E) + d_{\mathrm{TV}}(P,Q)\) via the
  now-proved (1.10)).
\end{proof}


\chapter{First passage, stabilization, and the main theorem}\label{chp:mainthm}

\section{The main theorem}

\begin{theorem}[Main theorem: Thm 1.3, Thm 1.6, quantitative Thm 3.1]\label{C6}
  \lean{tao_collatz, tao_collatz_quantitative}
  % C6s added 2026-07-15 (operator layer): the headline stubs discharge by `exact` from the
  % spine theorems, so the spine is C6's direct proof-dep.  The coarse pre-decomposition
  % edges (C7/C8/C9) are kept until the judge pass prunes; redundant, not wrong.
  \uses{C3, C7, C8, C9, C6s}
  \leanok
\textbf{Theorem 1.3} (the summit): for every \(f:\mathbb{N}\to\mathbb{R}\) with
\(f(N)\to\infty\), \(\mathrm{Colmin}(N)<f(N)\) for almost all \(N\) in logarithmic density.
It follows from the intermediate \textbf{Theorem 1.6} (almost all orbits attain almost bounded
values) which in turn follows from the \textbf{quantitative Theorem 3.1}: there exist
\(c,C>0\) so that for all \(N_0,x\ge 2\),
\[
  \mathrm{logProb}_{[1,x]}\bigl(\{N : \mathrm{Colmin}(N)\le N_0\}\bigr)
    \;\ge\; 1 - C/(\log N_0)^{c}.
\]
Thm~3.1 is obtained from the stabilization estimate Prop~1.11 (node~\ref{C9}) by iterating
over dyadic scales (§3).
\lapsrisk{0}{done}{100\%}
\end{theorem}
\begin{proof}
  \leanok
  \uses{C6s}
  🏔️ \textbf{THE SUMMIT} (2026-07-15, \texttt{8ccd1a2}/\texttt{9f17733}): both frozen
  \texttt{Statement.lean} stubs discharged by \texttt{exact} from the C6s spine — the only
  edit that file ever receives.  Kernel-checked axiom-clean host-side, 114 declarations
  (\texttt{blueprint\_audit} 2026-07-15, 0 drift, 0 false-green).  Judge ratification reads
  still owed on the C6 pins (ratify-on-pin flags stand); flips under the trust-the-treadmill
  ruling.
\end{proof}

% ---- C6 intermediates, pinned 2026-07-15 (directive step 2: the §3 reduction chain) ----
% All \notready pending judge ratification (rule 1: never self-\leanok).  Statements live in
% TaoCollatz/Sec3/Reduction.lean, sorried; numeric traps check14/check15.  The chain:
% stabilization (C9) => C6a => C6b => (with C6c) the two Statement.lean headlines via C6s.

\begin{theorem}[Thm 3.1, Syracuse forms]\label{C6a}
  \lean{tao_syracuse_quantitative_sum, tao_syracuse_quantitative}
  \uses{C1, C3, C9}
  \leanok
Paper p.16, both displays of the one claim ("\dots or equivalently \dots"):
\(\sum_{N\in 2\mathbb{N}+1\cap[1,x]:\ \mathrm{Syrmin}(N)>N_0} 1/N \ll \log x/\log^c N_0\),
and \(\mathbb{P}(\mathrm{Syrmin}(\mathbf{Log}(2\mathbb{N}+1\cap[1,x]))\le N_0) \ge
1-O(\log^{-c}N_0)\).  Obtained from Prop~1.11 (node~\ref{C9}) by the dyadic-scale
telescoping of \S3 (pp.17--18).
\lapsrisk{0}{done}{100\%}
\end{theorem}
\begin{proof}
  \leanok
  Kernel-clean, BOTH forms (\texttt{2c3e273}+; \texttt{blueprint\_audit} 2026-07-15 lists
  C6a on the MISSED-FLIP roll).  Telescope \(\to\) covering \(\to\) sum/probability forms.
  Trust-the-treadmill flip (operator layer); judge flip-back allowed.
\end{proof}

\begin{theorem}[Theorem 1.6]\label{C6b}
  \lean{tao_syracuse}
  \uses{C1, C3, C6a}
  \leanok
% Proof flipped same pass: went kernel-clean while the 2a01077 flip pass was being written.
Paper p.4: for \(f\) with \(f(N)\to\infty\), almost all odd \(N\) (log density on the odd
window) satisfy \(\mathrm{Syrmin}(N)<f(N)\).  From Thm~3.1 at \(N_0:=\tilde f(x)=
\inf_{N\ge x}f(N)\) (p.18).
\lapsrisk{0}{done}{100\%}
\end{theorem}
\begin{proof}
  \leanok
  Kernel-clean (\texttt{blueprint\_audit} 2026-07-15, MISSED-FLIP roll).  Trust-the-treadmill
  flip (operator layer); judge flip-back allowed.
\end{proof}

\begin{lemma}[The (1.2) odd-part reduction, both forms]\label{C6c}
  \lean{logSum_oddPart_pullback, almostAllPos_oddPart_of_almostAllOdd}
  \uses{C1, C3}
  \leanok
Worker-authored rendering of the one-paragraph reductions (p.5 and p.16 "in particular, by
(1.2)"): quantitatively, \(\sum_{N\le x:\ \mathrm{oddPart}(N)\in A} 1/N \le
2\sum_{M\in A\cap(2\mathbb{N}+1)\cap[1,x]} 1/M\) (geometric series over \(\nu_2\));
qualitatively, an almost-all-odd property pulls back along \(\mathrm{oddPart}\) to
almost-all on \(\mathbb{N}+1\).  Rests on the proved \(\mathrm{Colmin}=
\mathrm{Syrmin}\circ\mathrm{oddPart}\) (node~\ref{C1}).  Trap: check15 (constant 2 exact
and \(>1.8\)-tight, so a constant-1 mis-rendering fails).
\lapsrisk{0}{done}{100\%}
\end{lemma}
\begin{proof}
  \leanok
  Kernel-clean, both forms (\texttt{42fb97c} half 1, \texttt{3de7831} half 2;
  \texttt{blueprint\_audit} 2026-07-15).  Trust-the-treadmill flip; judge flip-back allowed.
\end{proof}

\begin{theorem}[Headline spine: Thm 1.3 and Thm 3.1-Colmin from the intermediates]\label{C6s}
  \lean{tao_collatz_spine, tao_collatz_quantitative_spine}
  \uses{C6a, C6b, C6c}
  \leanok
Statements byte-identical to the two frozen \texttt{Statement.lean} headlines; when these
close, the frozen sorries discharge by \texttt{exact} (the only edit
\texttt{Statement.lean} ever receives).  Quantitative route: C6a(sum form) + C6c +
harmonic-mass bounds on the full window.  Headline route: C6b at \(\tilde f\) + C6c +
\(\mathrm{oddPart}(N)\le N\).
\lapsrisk{0}{done}{100\%}
\end{theorem}
\begin{proof}
  \leanok
  Kernel-clean, both spines (\texttt{8ccd1a2}; \texttt{blueprint\_audit} 2026-07-15).
  Statements byte-identical to the frozen headlines, so the stubs discharged by
  \texttt{exact}.  Trust-the-treadmill flip; judge flip-back allowed.
\end{proof}

% ---- Ruling II successor (2026-07-17): the assembled explicit big-C layer.  Additive on
% the clean library; the retired CTao pin is NOT this node.  Numeric/structural trap: check27.
% \leanok set by JUDGE ratification 2026-07-17 (host-verified, NOT self-set by a lap):
% #print axioms {tao_collatz_quantitative_assembled, C_tao_assembled} = [propext,
% Classical.choice, Quot.sound] (no sorryAx); ExplicitnessClosure walked 209 project defs
% with no witness selector; big_c_cutoff_audit --complete green; check27 pass; library
% sorry census 0; statement read against BIG_C_EXPLICIT_BOUND_PLAN.md "Exact deliverable".
\begin{theorem}[Assembled fully-explicit Thm 3.1 (our augmentation)]\label{C6x}
  \lean{X_spine, C_tao_assembled, tao_collatz_quantitative_assembled}
  \uses{C6a, C6s}
  \leanok
Beyond the paper: Theorem 3.1 with BOTH slots closed --- the explicit exponent
\(c_{\mathrm{Tao}}\) and a closed (tower-valued, deliberately non-small) constant
\(C_{\mathrm{assembled}} = \max(C_{\mathrm{spine}}(X_{\mathrm{spine}}),
\log^{c_{\mathrm{Tao}}}2)\), where \(X_{\mathrm{spine}}\) is the X-chase closure of the
development's cutoff tree (no existential exponent, constant, or cutoff on the proof
path; explicitness contract per \texttt{BIG\_C\_EXPLICIT\_BOUND\_PLAN.md}, enforced by
\texttt{tools/ExplicitnessClosure.lean} + \texttt{big\_c\_cutoff\_audit.py --complete}).
\lapsrisk{0}{low}{100\%}
\end{theorem}
\begin{proof}
  \leanok
  Kernel-clean (axioms = the standard three, no \texttt{sorryAx}; host-verified
  2026-07-17 — receipt in the source comment above this node).  Delegation:
  \texttt{tao\_collatz\_quantitative\_spine\_atCX} at the closed cutoff, weakened to the
  exponent \(c_{\mathrm{Tao}}\) by monotonicity (\texttt{\_of\_le}); the \(N_0 = 2\)
  window is absorbed by the second \texttt{max} arm.  \texttt{BigCTower.lean} then proves
  the readable ceiling \(C_{\mathrm{assembled}} \le 10\uparrow\uparrow 63\)
  (\texttt{tenTower 62}), rendered on the trusted surface as \texttt{CTao} +
  \texttt{tao\_collatz\_quantitative\_fully\_explicit} (\texttt{Statement.lean}).
\end{proof}

\section{First passage (§5)}

\begin{definition}[First-passage data and the log-uniform window]\label{C7d}
  \lean{passes, passTime, passLoc, logWindow, logUnifOdd, alpha}
  \uses{C1}
  \leanok
The first-passage data: \(\mathrm{passes}(x,N)\) (\(=T_x(N)<\infty\), the orbit reaches
\(\le x\)), the passage time \(\mathrm{passTime}\), and the passage location
\(\mathrm{Pass}_x = \mathrm{passLoc}\) (with the convention \(\mathrm{Syr}^\infty := 1\));
plus the log-uniform law \(\mathbf{N}_y \equiv \mathbf{Log}(2\mathbb{N}+1 \cap [y,y^\alpha])\)
(\(\mathrm{logWindow}\), \(\mathrm{logUnifOdd}\)) and the exponent \(\alpha = 1.001\) (paper (1.18)).
\lapsrisk{0}{low}{100\%}
\end{definition}

% ONE NODE, ONE CLAIM (see blueprint_rules.md).  Until 2026-07-14 this lemma and the definition
% node above were a SINGLE node: a \begin{lemma} whose \lean{} named three DEFS and which carried
% \leanok.  The defs were done, so the node rendered GREEN and was rated "low / 5--10 laps" --
% while its actual content, the estimate (1.19), was not in Lean at all.  Both halves of the
% display were reporting on the half that was already finished.  Split at judge pass 29.
\begin{lemma}[First-passage non-escape; (1.19)]\label{C7}
  \lean{first_passage_nonescape}
  \uses{C5, S2, C7d}
  \leanok
Paper (1.19): a log-uniformly chosen odd \(N_y\) in the window \([y,y^\alpha]\) fails ever to
descend to \(\le x\) only with small probability,
\[
  \mathbb{P}\bigl(T_x(N_y)=\infty\bigr) \ll x^{-c}.
\]
Tao's route (\S5 pp.20--21) runs entirely over proved machinery except for one new brick: the
\textbf{integral test} \(d_{\mathrm{TV}}(\mathbf{N}_y \bmod 2^{n'}, \mathrm{Unif}) \ll 2^{-n'}\),
which is exactly the hypothesis Proposition~1.9 (node C5) takes.  Given it, Prop~1.9 supplies
(5.4), Lemma~2.2 (node S3, two-sided) supplies the lower-tail bound (5.5)
\(\mathbb{P}(|\vec a^{(n_0)}(\mathbf{N}_y)| \le 1.9 n_0) \ll x^{-c}\), and an explicit descent
computation gives \(\mathrm{Syr}^{n_0}(\mathbf{N}_y) = O(x^{0.99}) \le x\) on the complement,
with \(n_0 := \lfloor \log x / (10\log 2)\rfloor\) (5.1).
\lapsrisk{10--18}{medium}{75\%}
\end{lemma}

\begin{proof}
  \leanok
  Ratified and flipped at judge pass 30 (2026-07-14): the theorem is kernel-clean
  (\(\mathrm{propext}, \mathrm{Classical.choice}, \mathrm{Quot.sound}\)), clearing the pass-29
  missed flip.  The route closed as designed --- the integral test discharges Proposition~1.9's
  \(d_{\mathrm{TV}}\) hypothesis (5.4), the two-sided Lemma~2.2 tail gives (5.5), and an explicit
  descent yields \(\mathrm{Syr}^{n_0}(\mathbf N_y) = O(x^{0.99}) \le x\).
\end{proof}

\begin{proposition}[Approximate first-passage formula; Prop 5.2]\label{C8}
  \lean{first_passage_approx}
  \uses{C2, C5, C7}
  \leanok
\textbf{Proposition 5.2}, the approximate formula (paper (5.8)) for the law of the
first-passage location, built from the bookkeeping events
\(\mathcal{A}^{(n')}\) (5.11), \(E'\) (5.10), and \(I_y\) (5.9), through the
\(B_{n,y}\) equivalence chain relating the passage location to the affine images of the input
window.  This is where the Prop~1.9 geometric approximation is fed into the concrete window
statistics.
% STATEMENT RATIFIED judge pass 30 (2026-07-14) as RATIFY-C8-v2: first_passage_approx +
% approxMainTerm read VERBATIM against Prop 5.2 (5.8)/(5.9)/(5.10)/(5.11)/(5.18)/(5.19) on PDF
% pp.22--25.  The v1 pin (N-truncating Aff, unguarded) over-counted (5.8) by a super-polylog
% factor and made truncation_error_bound FALSE; v2 guards the pushforward by the EXACT affine
% relation 3^{n-m0}*N + fnat = M*2^{|a|} (= the (5.18) congruence + (5.19) value), so Lemma 2.1
% makes the reindex exact (approxMainTerm_eq_steppedMid, PROVED axiom-clean).  The statement
% \leanok is RATIFY-C8-v2 (judge pass 30); the proof \leanok below is the 2026-07-15 flip.
\lapsrisk{0}{done}{100\%}
\end{proposition}
\begin{proof}
  \leanok
  \uses{C7}
  🏆 \textbf{COMPLETE} (2026-07-15 overnight): the (5.16) window term closed via the integral
  test over C7's \texttt{classMass}/\texttt{windowMass} machinery; the (5.17) stepback closed
  by the forward/reverse-leg split with the early-return event proved \emph{empty}
  (\texttt{0bea9d1}).  Kernel-checked axiom-clean by \texttt{blueprint\_audit} 2026-07-15
  (MISSED FLIP list, twice).  Statement pre-ratified (pass 30), so this flip is the mechanical
  half — done by the operator layer per the DIRECTION.md operator note.
\end{proof}

\begin{proposition}[Stabilization across scales; Prop 1.11]\label{C9}
  \lean{stabilization}
  \uses{C8, C10, C9A, C9B1, C9B2}
  \leanok
\textbf{Proposition 1.11}, the spine's key input.  With \(\alpha=1.001\) (paper (1.18)):
there are \(c,C,x_0>0\) so that for \(x\ge x_0\), the first-passage location laws seen from two
successive scales \(x^{\alpha}\) and \(x^{\alpha^2}\) agree up to
\[
  d_{\mathrm{TV}}\Bigl(
      \bigl(\mathrm{logUnifOdd}_{[x^{\alpha},\,x^{\alpha^2}]}\bigr)_*\mathrm{Pass}_{\lfloor x\rfloor},\
      \bigl(\mathrm{logUnifOdd}_{[x^{\alpha^2},\,x^{\alpha^3}]}\bigr)_*\mathrm{Pass}_{\lfloor x\rfloor}
    \Bigr) \le C(\log x)^{-c},
\]
together with the non-escape control on each window.  Proved via Lemma~5.3
(\(c_n(X)\ll 1\)) and (5.18)--(5.21), applying the fine-scale mixing Prop~1.14
(node~\ref{C10}) at a base scale \(m_0\).
% TODO-PRECISE: window endpoints (nested \(\alpha\)-powers, RATIFY-3) to be ratified.
\lapsrisk{0}{done}{100\%}
\end{proposition}
\begin{proof}
  \leanok
  \uses{C8, C10, C9A, C9B1, C9B2}
  🏆 \textbf{COMPLETE} (2026-07-15, \texttt{9e5604b}): assembly + every leaf closed;
  \texttt{stabilization} kernel-checked axiom-clean by \texttt{blueprint\_audit} (MISSED FLIP
  list).  Flipped under the trust-the-treadmill ruling (DIRECTION.md operator note); judge
  flip-back allowed.
\end{proof}

% --- C9 sub-nodes B1/B2 (registered retroactively 2026-07-15, operator layer; DIRECTION.md
% operator note).  The pass-30 refresh mandated decomposing harmonic_to_Z (5.20) into B1/B2;
% the worker pinned + (partly) proved them, but the names lived only in handoffs/commits while
% mimicking node IDs.  Registered here so every node-shaped name resolves on the map.  Both
% \notready pending judge ratification (rule 1: never self-\leanok) -- the statements exist and
% compile (Sec5/Stabilization.lean), so orange UNDERSTATES progress, the safe direction.

\begin{lemma}[C9 leaf B1 -- geomHalf\(\to\)syracZ reindex; (5.20) left half]\label{C9B1}
  \lean{perNHarmonic_eq_harmZfine_approx}
  \leanok
\(|\mathrm{perNHarmonic}\,x\,E\,n - \mathrm{harmZfine}\,x\,E\,n| \le C(\log x)^{-c}\) on
\(I_y\): the iid-geomHalf \(2^{-\mathrm{pre}\,\bar a}\) mass over good, affine-solvable tuples
agrees with the exact \(\mathrm{Syrac}(\mathbb{Z}/3^{n-m_0}\mathbb{Z})\) mass.  Both ribs
(\texttt{syracZ\_sub\_perNGoodMass\_bound}; the \texttt{solvable\_iff\_fmapZ} pointwise bridge)
are proved; the assembly is L\(^1\times\)L\(^\infty\) H\"older against \texttt{cn\_bound}.
Does NOT consume C10.
\lapsrisk{0}{done}{100\%}
\end{lemma}
\begin{proof}
  \leanok
  Kernel-clean (\texttt{44fcd94}; \texttt{blueprint\_audit} 2026-07-15).  Trust-the-treadmill
  flip (operator layer); judge flip-back allowed.
\end{proof}

\begin{lemma}[C9 leaf B2 -- the C10 seam; (5.20) scale bridge]\label{C9B2}
  \lean{harmZfine_to_mainZ}
  \uses{C10}
  \leanok
\(|\mathrm{harmZfine}\,x\,E\,n - \mathrm{mainZ}\,x\,E| \le C(\log x)^{-c}\):
L\(^1\times\)L\(^\infty\) H\"older, \(\sup_X c_n(X) \le C\log^{0.7}x\) (\texttt{cn\_bound})
against \(\mathrm{osc}\,m_0\,(n-m_0) \le C' m_0^{-A}\) via Prop~1.14 (node~\ref{C10}) at
\(A > 0.7 + c\) -- the SOLE consumption of C10 inside C9.
\lapsrisk{0}{done}{100\%}
\end{lemma}
\begin{proof}
  \leanok
  Kernel-clean (\texttt{ece1f36}; \texttt{blueprint\_audit} 2026-07-15).  Trust-the-treadmill
  flip (operator layer); judge flip-back allowed.
\end{proof}

\begin{lemma}[C9 leaf A -- (5.19) harmonic reduction of the per-\(n\) term]\label{C9A}
  \lean{perNTerm_harmonic_approx}
  \leanok
% Registered 2026-07-15 (worker, vocabulary rail from the operator note): the leaf-A
% decomposition names coined this run.
\(|\mathrm{perNTerm}\,x\,E\,y\,n - \mathrm{perNHarmonic}\,x\,E\,n/\mathrm{norm}| \le
C(\log x)^{-c}/\mathrm{norm}\), \(\mathrm{norm} = \tfrac{\alpha-1}{2}\log y\): the affine mass
collapses to the single point \(N^* = (M\,2^{\mathrm{pre}\,\bar a} - \mathrm{fnat})/3^{n-m_0}\)
(\texttt{perNTerm\_pointmass}), which is odd (\texttt{Nstar\_odd}, via \texttt{fnat\_odd}) and
lies on the window (\texttt{Nstar\_mem\_logWindow}, both PROVED this run), so
\(\mathbb{P} = (N^*)^{-1}/D_y\); then \((N^*)^{-1} \approx 3^{n-m_0}/(M\,2^{\mathrm{pre}})\)
(relative \(O(x^{-c})\)) and \(1/D_y \approx 1/\mathrm{norm}\)
(\texttt{windowMass\_estimate}/\texttt{windowMass\_ge\_clog}), with relative errors made
absolute by \(\mathrm{perNHarmonic} \le C\log^{0.7}x\) (\texttt{perNHarmonic\_le}, PROVED).
Does NOT consume C10.
\lapsrisk{0}{done}{100\%}
\end{lemma}
\begin{proof}
  \leanok
  Kernel-clean (\texttt{60bb490}; \texttt{blueprint\_audit} 2026-07-15).  Trust-the-treadmill
  flip (operator layer); judge flip-back allowed.
\end{proof}


\chapter{Fine-scale mixing (§6)}\label{chp:mixing}

\begin{proposition}[Fine-scale mixing; Prop 1.14 \(\Leftarrow\) Prop 1.17]\label{C10}
  \lean{fine_scale_mixing}
  \uses{C4, S2, S4, X11}
  \leanok
\textbf{Proposition 1.14}: the Syracuse law is asymptotically equidistributed at fine scales,
\[
  \forall A>0,\ \exists C>0,\ \forall\, 1\le m\le n:\quad
  \mathrm{Osc}_{m,n}\bigl(Y\mapsto (\texttt{syracZ}\,n)(Y)\bigr) \le C\,m^{-A}.
\]
Deduced from the \textbf{character-sum decay} Prop~1.17 (node~\ref{X11}, the head of the §7
crux ``X-chain'') by Plancherel on \(\mathbb{Z}/3^n\mathbb{Z}\).  The §6 machinery around it:
Lemma~6.2 (\(F_n\) injective), Cor~6.3 (\(3\)-adic separation of the offsets), the event \(E\)
(6.2), the stopping time \(k\), and the Plancherel step.
\lapsrisk{0}{done}{100\%}
\end{proposition}
\begin{proof}
  \leanok
  \uses{X11}
  🏆 \textbf{COMPLETE and judge-verified} (2026-07-14, HEAD \texttt{49b32c7}):
  \texttt{\#print axioms fine\_scale\_mixing} \(= [\texttt{propext}, \texttt{Classical.choice},
  \texttt{Quot.sound}]\), no \texttt{sorryAx}.  The node collapsed to a single tail estimate via two
  machine-checked identities — \texttt{mainHigh\_eq\_restrictedDensity} (\texttt{mainHigh} \emph{is}
  the Syracuse pushforward restricted to \texttt{mainEvent}) and
  \texttt{sum\_abs\_syracZ\_sub\_mainHigh\_eq} (\(\sum_Y|\texttt{syracZ}-\texttt{mainHigh}| =
  \mathbb{P}(\neg\texttt{mainEvent})\), an \emph{equality}) — after which \(\texttt{globalGood}
  \subseteq \texttt{mainEvent}\) and the (6.3) union bound over the three deviation tails
  (\(g_1,g_2,g_3\)) closed \texttt{error\_l1\_high\_bound}, itself axiom-clean.  \texttt{hbudget}
  discharged from the tight window (\texttt{lRange\_hbudget}); the \(A'\)-absorption at \(C_A=30\)
  \emph{shown}, not asserted (\texttt{osc\_mainHigh\_bound}).
\end{proof}


\chapter{The crux: character decay via renewal versus triangles (§7)}\label{chp:crux}

The entire descent bottoms out at Prop~1.17.  §7 proves it by studying the phase functions
\(\theta(j,l)\) attached to the paired valuations, splitting the index set into ``white''
points (where the character sum contracts) and ``black'' points (where it does not), showing
the black set is a union of separated triangles, and running a renewal process that is forced
to hit many white points --- unless it is trapped in a large triangle, which is rare.

\section{Setup and the phase functions}

\begin{definition}[§7 setup: character, pairing, factorization]\label{X1}
  \lean{fCond, xArg, cexpect_pairing}
  \uses{C4, S2}
  \leanok
The character \(\chi\) (paper (7.1)) and the reduction of Prop~1.17 to a bound on
\(\mathbb{E}\,\chi\) through the reversed form (1.26); the \textbf{pairing}
\(b_j = a_{2j-1}+a_{2j}\sim\mathrm{Pascal}\) that groups the geometric valuations two at a
time; and the conditional \textbf{factorization} (7.4)/(7.5) of the Fourier coefficient into a
product of per-pair factors \(f\) (with the low-order piece \(g\) split off).
\emph{Pinned + ratified 2026-07-12 (judge pass 10, lap 53, paper pp.33--35)}:
\texttt{cexpect\_pairing} is the (7.5) bound verbatim --- LHS = the (7.2) reversed character
sum over \(\mathrm{Geom}(2)^n\), RHS = the \(\mathrm{Pascal}^{\lfloor n/2\rfloor}\)
expectation of \(\prod_j \|f(x_j,b_j)\|\), the \(|g|\le 1\) leftover drop built in, and no
\(3\nmid\xi\) hypothesis needed (pure algebra, a harmless strengthening).  \texttt{fCond} is
(7.4) in concrete uniform-pair form (junk value \(0\) off Pascal's support, documented);
\texttt{xArg} carries the 0-based \(j\)-shift (RATIFY-4).  Character algebra, (7.6), and
Lemma~7.2's exact value \(\|f(x,3)\| = |\cos\pi\theta|\) are \emph{proved};
\texttt{key\_fourier\_decay} (Prop~7.1) moved drift-free from \texttt{Holding.lean} and is
now a \emph{theorem} from the pairing + damping + Prop~7.3, and \texttt{charFn\_decay}
(Prop~1.17) is derived across the (1.26) seam.
\textbf{COMPLETE (lap 53; judge pass 11, 2026-07-12)}: \texttt{cexpect\_pairing} PROVED via
the generalized pair-peel \texttt{cexpect\_pairing\_gen} (statement drift-free) and
judge-verified \([\mathrm{propext}, \mathrm{Classical.choice}, \mathrm{Quot.sound}]\);
Prop~1.17's remaining trust runs only through the Prop~7.8 chain
(\texttt{white\_cos\_bound} closed lap 54, judge pass 12).
\end{definition}

\begin{lemma}[Phase identities and white-point cancellation; Lemma 7.2]\label{X2}
  \lean{sfrac, θq, θq_succ_j, θq_pred_l, black, fCond_three_norm, white_cos_bound}
  \uses{X1}
  \leanok
The phase \(\theta(j,l)\) (paper (7.7)/(7.8)), a signed fractional part of
\(\xi\cdot 3^{2j-2}\cdot 2^{\,1-l}/3^n\), and its two scaling identities (7.13)/(7.14): there
are integers \(k\) with
\[
  \theta(j{+}1,l) = 9\,\theta(j,l) + k, \qquad \theta(j,l{-}1) = 2\,\theta(j,l) + k .
\]
The black/white dichotomy (7.9) with \(\varepsilon = 10^{-4}\): a point is \emph{black} when
\(|\theta(j,l)|\le\varepsilon\), \emph{white} otherwise.  \textbf{Lemma 7.2} (white-point
cancellation): at a white point the per-pair factor contracts,
\(|f| \le |\cos(\pi\theta)| \le \exp(-\varepsilon^3)\).
\end{lemma}
\begin{proof}
  \leanok
COMPLETE (campaign laps 53--54; judge pass 12, 2026-07-12).  Both halves of Lemma 7.2:
\texttt{fCond\_three\_norm} (exact value at the paired point, lap 53) and
\texttt{white\_cos\_bound} (the Taylor half, lap 54): white \(\Rightarrow \varepsilon <
|\theta| \le 1/2 \Rightarrow |\cos\pi\theta| \le 1 - 2\theta^2 \le 1 - 2\varepsilon^2 \le
e^{-\varepsilon^3}\) (via mathlib's \texttt{Real.cos\_le\_one\_sub\_mul\_cos\_sq}).
Judge-run \texttt{\#print axioms} (2026-07-12): [propext, Classical.choice, Quot.sound]
on both; the damping consumer \texttt{prod\_fCond\_le\_damping} closed with it.
\end{proof}

\section{The triangle structure of the black set}

\begin{lemma}[Black set = separated triangles; Lemma 7.4]\label{X3}
  \lean{triangle, black_structure}
  \uses{X2}
  \leanok
\textbf{Risk kernel 2.}  With a triangle (paper (7.11)) the set
\(\{(j,l) : j_0\le j,\ l\le l_0,\ (j-j_0)\log 9 + (l_0-l)\log 2 \le s\}\), Lemma~7.4 asserts
the black set inside the strip is a disjoint union of triangles that are (i) confined to the
strip \(j+1\le \lfloor n/2\rfloor - \tfrac{1}{10}\log(1/\varepsilon)\), and (ii) pairwise
Euclidean-separated by at least \(\tfrac{1}{10}\log(1/\varepsilon)\):
\[
  \{(j,l) : j{+}1\le \lfloor n/2\rfloor,\ \text{black}(j,l)\}
    \;=\; \bigsqcup_{t\in T} \mathrm{triangle}(t).
\]
Proved through the phase identities (7.12)--(7.15), the weakly-black claims (i)--(iii), the
\(l^*/j^*\) construction, and Claim~(\(*\)) Cases~1--3.
% TODO-PRECISE: exact conjunction spelling (RATIFY-5) is implementer's choice.
% Risk kernel 2 CLOSED 2026-07-10 (box laps 6--10, est. was 15--30 laps / 75%).
\end{lemma}
\begin{proof}
  \leanok
  \texttt{black\_structure} in \texttt{Sec7/Triangles.lean}, via the exact fibre identity
  \(\theta(j,l) = 9^{j-j^*}2^{l^*-l}\theta^*\) (sharper than the paper's (7.18) inequality;
  at \(\varepsilon = 10^{-4}\) the separation conjunct reduces to lattice disjointness).
  Judge-verified \texttt{\#print axioms}: \texttt{[propext, Classical.choice, Quot.sound]}
  (2026-07-10, host \texttt{lake env lean}).
  \emph{Pass-18 note (route escalation):} that lattice reduction means the separation
  clause carries no content beyond disjointness at the frozen \(\varepsilon\) --- sound
  for THIS statement, but the (7.50) white-exit ring consumes real separation, so any
  D4 \(\varepsilon\)-shrink reopens this clause with a genuine Euclidean obligation
  (pp.39--41; \texttt{sep\_const\_sq\_le\_one} dies by design once
  \(\mathrm{sep}^2 > 1\)).  Node status unaffected.
  \emph{Pass-23 (2026-07-13): the obligation is DISCHARGED FOR REAL.}  An external
  Codex session executed the D4 change (\(\varepsilon = 10^{-90}\), so
  \(\mathrm{sep} = 9\ln 10 \approx 20.7\), \(20 < \mathrm{sep} < 26\) proved) and
  formalized Lemma 7.4's Claim~(\(*\)) Cases 1--3 as genuine Euclidean-separation
  content; \texttt{sep\_const\_sq\_le\_one} is deleted as designed.  The statement
  gained one \emph{additive} conjunct (every family triangle is the
  \texttt{cornerTriple} of a black point --- the paper's own construction,
  re-ratified) and is judge re-verified clean at the new \(\varepsilon\)
  (2026-07-13 dated run).
\end{proof}

\section{The holding process and its finitized renewal recursion (D6)}

\begin{proposition}[Holding law, the \(Q\) recursion, and Prop 7.3]\label{X4}
  \lean{hold, Q_rec, Q_boundary, renewal_white_encounters}
  \uses{S2, X2}
  \leanok
The \textbf{holding law} \(\mathrm{Hold}\) on \(\mathbb{N}\times\mathbb{Z}\) (§7.3), whose first
coordinate is always \(\ge 1\), governing how the renewal walk steps between pairs.  Design
decision D6 replaces the paper's infinite renewal product (7.34) by the \emph{defined}
downward recursion \(Q:\mathbb{N}\times\mathbb{Z}\to\mathbb{R}\) (paper (7.35)),
\[
  Q(j,l) = 1 \ \ (j>\lfloor n/2\rfloor); \qquad
  Q(j,l) = e^{-\varepsilon^3\mathbf 1_W(j,l)}\cdot\!\!\sum_{d}\!(\mathrm{Hold}\,d)\,Q\bigl((j,l)+d\bigr)
  \ \ (j\le\lfloor n/2\rfloor),
\]
well-founded on \(\lfloor n/2\rfloor - j\) (Hold's first coordinate is \(\ge 1\)), with
\(0\le Q\le 1\).  The bridge (7.28)/(7.34)--(7.36) makes
\(\mathbb{E}\,Q(\mathrm{Hold})\ll_A n^{-A}\) equivalent to \textbf{Proposition 7.3}
(the renewal white-encounter count, in the finite-vector form over \(\mathrm{Pascal}^{\otimes\lfloor n/2\rfloor}\)).
This node exists to validate the D6 finitization.
\lapsrisk{8--15}{medium}{80\%}
\end{proposition}

\begin{proof}
  \leanok
  PROVED.  \emph{Judge pass 27 (2026-07-14), dated \texttt{\#print axioms} at pin
  \texttt{8505bd4}}: \texttt{Q\_rec}, \texttt{Q\_boundary},
  \texttt{renewal\_white\_encounters} all \(=\)
  \([\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}]\) --- the trust base
  only.  The D6 downward recursion is \emph{defined} (well-founded on
  \(\lfloor n/2\rfloor - j\)), so the (7.28)/(7.34)--(7.36) bridge is a theorem rather than a
  construction obligation.
\end{proof}

\begin{lemma}[Holding-law basics; Lemma 7.6]\label{X5}
  \lean{hold_mean_fst, hold_mean_snd, hold_aperiodic}
  \uses{X4}
  \leanok
\textbf{Lemma 7.6} (paper p.42): the explicit distribution of \(\mathrm{Hold}\), its
exponential tail, aperiodicity, and mean vector \((4,16)\).  These are the renewal-theory
inputs (drift and spread) for the first-passage analysis of the walk.  Clause map:
\emph{exponential tail} + the ``in particular'' Lemma 2.2 conclusion are the S3 engine
(\texttt{hold\_tail\_bound}/\texttt{hold\_local\_bound}, proved directly by the paper's own
Chernoff route (7.29)--(7.30)); \emph{mean \((4,16)\)} = \texttt{hold\_mean\_fst/snd}
(direct sum over the explicit construction, \(\mathbb{E}\mathrm{Pascal}' = 13/3\) per
(7.29)); \emph{aperiodicity} = \texttt{hold\_aperiodic} (support \(\subseteq x + H\)
forces \(H = \top\), the paper's own witnesses \((1,3)\), \((1,3)+(1,b)\)).
\end{lemma}
\begin{proof}
  \leanok
COMPLETE (campaign lap 54; judge pass 12, 2026-07-12).  Statements ratified vs paper p.42
(previously an unread front): the (7.29) mass function is character-exact, the means follow
the paper's own arithmetic, and the coset formulation is faithful.  Judge-run
\texttt{\#print axioms} (2026-07-12): [propext, Classical.choice, Quot.sound] on all
fifteen \texttt{HoldBasics} declarations.
\end{proof}

\begin{lemma}[First-passage location of the holding walk; Lemma 7.7]\label{X6}
  \lean{fpDist, fpDist_location_bound}
  \uses{S3, X5}
  \leanok
\textbf{Lemma 7.7} (paper p.43): the first-passage endpoint mass satisfies
\(\mathbb{P}(\mathbf{v}_{[1,\mathbf{k}]}=(j,l)) \ll e^{-c(l-s)}(1+s)^{-1/2}
G_{1+s}(c(j-s/4))\).  D6 form: \texttt{fpDist} (budget-recursive endpoint PMF, support
facts proved) and \texttt{fpDist\_location\_bound}, judge-ratified vs the p.43 display
2026-07-10 (unconditional in \(l\) since the support fact kills \(l\le s\)).
Uses the local 2-D bound (node~\ref{S3}) applied to the holding sums.
\end{lemma}
\begin{proof}
  \leanok
COMPLETE (campaign laps 46--50; judge pass 7, 2026-07-12).  Route: renewal decomposition
\texttt{fpDist\_le\_renewal\_conv} (budget induction --- monotone height kills the barrier
condition) + \texttt{renewalMass\_bound} (sum of Lemma 2.2(i) in \(k\), three regions) +
last-step assembly (\texttt{hold\_step\_bound}, Gaussian\(\times\)exponential convolution
\texttt{conv\_Gweight\_exp}, recentring shift, AP/\(\ell^1\) envelopes).  Statement verified
character-identical to the pass-3 ratified form.  Judge-run \texttt{\#print axioms}
(2026-07-12): [propext, Classical.choice, Quot.sound] on
\texttt{fpDist\_location\_bound}, \texttt{renewalMass\_bound}, \texttt{hold\_step\_bound},
\texttt{conv\_Gweight\_exp}.
\end{proof}

\section{The three cases of Prop 7.8}

\begin{proposition}[Monotone functional \(Q_m\); Prop 7.8 skeleton; Case 1]\label{X7}
  \lean{Qm, prop_7_8, Q_white_case1, Q_black_edge, Q_polynomial_decay}
  \uses{X4}
  \leanok
The truncated functional \(Q_m\) (paper (7.38)) and the skeleton of \textbf{Proposition 7.8}
(the quantitative white-encounter bound), together with \textbf{Case 1} (paper (7.42)/(7.43)):
when the walk starts at a white point, the \(e^{-\varepsilon^3}\) contraction from Lemma~7.2
(node~\ref{X2}) already delivers the required decay.  Judge-ratified 2026-07-10 vs pp.45--46:
(7.38)\(\to\)\texttt{Qm}, (7.40)\(\to\)\texttt{prop\_7\_8}, (7.43)\(\to\)\texttt{Q\_white\_case1}
(verbatim incl.\ \(e^{-\varepsilon^3/2}\)), (7.37)\(\to\)\texttt{Q\_polynomial\_decay}.  Case 1
is \emph{proved}; the skeleton delegates the black edge to \texttt{Q\_black\_edge}
(nodes~\ref{X8}/\ref{X10}), whose Case-2/Case-3 split over the Lemma 7.4 triangle family
is now itself PROVED (judge pass 6) --- \texttt{prop\_7\_8} is fully conditional on the
two case statements alone.
\lapsrisk{4--8}{low}{85\%}
\end{proposition}

\begin{proof}
  \leanok
  PROVED --- \textbf{\S7 monotonicity is closed}.  \emph{Judge pass 27 (2026-07-14), dated
  \texttt{\#print axioms} at pin \texttt{8505bd4}}: \texttt{prop\_7\_8},
  \texttt{Q\_white\_case1}, \texttt{Q\_black\_edge}, \texttt{Q\_polynomial\_decay} all \(=\)
  \([\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}]\).  \texttt{prop\_7\_8}
  no longer carries \texttt{sorryAx}: Case 1 was already proved, and the black edge is now
  discharged through both branches --- Case 2 (node~\ref{X8}) and Case 3 (node~\ref{X11}).
\end{proof}

\begin{lemma}[Case 2: shallow inside a triangle]\label{X8}
  \lean{TriangleFamily, Q_black_edge_case2, Q_fp_endpoint_le, fpDist_edgeWeight_le, fpDist_white_exit, budget_le_of_mem_triangle}
  \uses{X3, X6, X7}
  \leanok
\textbf{Risk kernel 3a.}  \textbf{Case 2} (paper (7.44)--(7.51), pp.46--48): the walk starts
shallow inside a triangle (budget \(s\le m/\log^2 m\)); the analysis produces the \(\gg 1\)
white-exit bound (7.50)/(7.51) showing the walk leaves the triangle through a white in-strip
point with absolute probability \(p_0>0\).  Combines the triangle geometry (node~\ref{X3}),
the first-passage location law (node~\ref{X6}), and the Case-1 contraction (node~\ref{X7}).
Statements judge-ratified 2026-07-12 vs pp.46--48: the (7.46) endpoint step
(subtraction-form damping, PROVED) and the (7.52) budget bound (PROVED, with documented
lattice slack \(m+2\) for \(m\)); the two open kernels are the (7.48)/(7.49) weight
degradation and the (7.50)/(7.51) white-exit mass --- both consume node~\ref{X6}; the
(7.47)-split assembly is mechanical bookkeeping once they land.  The Lean white-exit form
strengthens the paper's ``white'' to ``white and in-strip'', which the damping mechanism
requires.
\lapsrisk{8--16}{medium --- risk kernel 3a, statements pinned}{75\% (2026-07-12 judge re-rate: endpoint step + budget PROVED axiom-clean; white-exit $\approx 0.99$ Monte Carlo, $\varepsilon$-sites verified 2026-07-10)}
\end{lemma}

\begin{proof}
  \leanok
  PROVED --- \textbf{Case 2 is closed} (the twelfth verified node).  \emph{Judge pass 26
  (2026-07-14), re-verified pass 27 at pin \texttt{8505bd4}}: \texttt{Q\_black\_edge\_case2},
  \texttt{fpDist\_white\_exit}, \texttt{fpDist\_edgeWeight\_le},
  \texttt{Q\_fp\_endpoint\_le}, \texttt{budget\_le\_of\_mem\_triangle} all \(=\)
  \([\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}]\).  Both kernels that
  were open at pin time landed: the (7.48)/(7.49) weight degradation
  (\texttt{fpDist\_edgeWeight\_le}, via the first-coordinate MGF engine) and the
  (7.50)/(7.51) white-exit mass (\texttt{fpDist\_white\_exit}, by kernel-merge relocation into
  \texttt{BlackEdgeQ.lean}).
\end{proof}

\begin{lemma}[Many triangles \(\Rightarrow\) many white points; Lemma 7.9]\label{X9}
  \lean{many_triangles_white, EncState, encStep, encVal, encExpect, encChainX, encounter_vertex_bound, fpDist_white_exit_deep}
  \uses{X4, X8}
  \leanok
\textbf{Re-pinned DEPTH-GATED + re-ratified (judge pass 14, 2026-07-12).}  The lap-55
truth challenge (judge-concurred: adversarial edge-strip families falsify the all-starts
\(e^{2\varepsilon}\) --- literature hole \#6, sibling of \#5; the paper's own Lemma 7.9
inherits it through ``(7.59) by repeating (7.51)'') is resolved by the box's preferred
fix: \texttt{encStep} gains a gate \(g\) (encounters count only at depth \(\ge g\) from
the strip edge; \(g = 0\) recovers the prior encoding definitionally), and
\texttt{many\_triangles\_white} binds \(g\) existentially.  Consumer-safe: Case 3's
(7.54) surviving branch keeps every (7.67) encounter at depth \(\ge 0.1m \ge g\) once
\(C_{A,\varepsilon} \ge 10g\) --- \textbf{judge-verified vs pp.48--49 + 54--56 (pass 15)},
along with the \(R := \lfloor A^2/\varepsilon^4\rfloor\)-after-\(\varepsilon\) choice and
the Markov \(-O(A)\) slack that absorbs the \(e^{2\varepsilon}\) constant.  Remaining
obligation: the proof's \(\varepsilon_0\) must be \(\ge\) the dichotomy constant
\(10^{-4}\) (the consumer instantiates there).  Both deviations carry full docstrings
with paper anchors.
\textbf{Lemma 7.9} (paper (7.57), pp.50--51): a walk crossing many triangles must encounter
many white points --- the banked white damping beats the \(\varepsilon\cdot\min(r,R)\) credit.
\emph{Statement pinned + ratified 2026-07-12 (judge pass 8, lap 52; re-ratified pass 9 after
the constant correction)}: the stopping-time data \((t_i, \Delta_i, r)\) is a left fold
(\texttt{EncState}/\texttt{encStep}: position, clearing barrier, encounter count, running +
banked white counts), and \texttt{many\_triangles\_white} states
\(\texttt{encExpect} \le e^{2\varepsilon}\) for all horizons \(T\), all \(R \ge 1\), all
\(\varepsilon \le \varepsilon_0\).
\textbf{Deviation from the paper (lap 52 route finding, judge-CONFIRMED pass 9):} the paper's
p.51 conditional-expectation display banks white damping through the full first block
(\(\sum_{p\le k_1}\)), but on stopped chains the true (7.57) sum ends at \(t_1 < k_1\) --- and
the discrepancy is typically the \(k_1\) exit-whiteness itself, the very term (7.59) cashes.
The display under-states the true conditional expectation, so the derivation is unsound as
written; the corrected pay-from-the-previous-block ledger has sharp toy value
\(e^{\varepsilon}p_0/(1-(1-p_0)e^{\varepsilon}) \approx e^{\varepsilon/p_0} > e^{\varepsilon}\),
hence the pin's \(e^{2\varepsilon}\) (valid for \(p_0 \ge 1/2\); the p.55 consumer is Markov
with \(R\) chosen after \(\varepsilon\), so absolute exponent constants wash out).
\emph{Riders}: (a) verify at node~\ref{X11} that p.55 chooses \(R\) after \(\varepsilon\);
(b) the \(2\varepsilon\) constant commits the proof to certifying \(p_0 > 1/2\) --- now
\emph{embodied} in the pinned external input \texttt{fpDist\_white\_exit\_deep} (judge pass
12, ratified vs the p.51 (7.59) reduction: identical to node~\ref{X8}'s Case-2 twin minus
the budget hypothesis, which the paper's ``repeating the proof of (7.51)'' at arbitrary
triangle points licenses, with the (7.52) budget \(s = O(m)\) free from membership; the
\(p_0 > 1/2\) mass demand is the corrected ledger's documented deviation from the paper's
\(\gg 1\), Monte Carlo \(\approx 0.99\)).
Other ratified deltas (pass 8): finite horizon \(\forall T\) (D6; T1 did NOT fire);
\(\exists\varepsilon_0\)-family; index shift matching the Case-2 forms.  Proved so far:
geometric prerequisites, head-peel \texttt{encExpect\_succ}, block bridge
\texttt{encExpect\_block\_le}, coupling \texttt{encExpect\_anti} (pass 9), and the
corrected per-block ledger's closed-form core (lap 54, pass 12): the chain value
\texttt{encChainX} \(= p_0/(1-(1-p_0)e^{\varepsilon})\) with fixed-point identity, the cap
\(X \le e^{\varepsilon}\) (whence \(e^{\varepsilon}X \le e^{2\varepsilon}\), the (7.57)
constant), and the four-mass LP vertex analysis \texttt{encounter\_vertex\_bound} (maximum
at \((0, 1-p_0)\) = the fixed point) --- all axiom-checked.  \textbf{Lap 55 (judge pass
16): the \(Y/Z\) induction is CLOSED} --- \texttt{many\_triangles\_white} is proved with
sorryAx trail exactly \(\{\texttt{fpDist\_white\_exit\_deep}\}\) (machine-checked); the
kernel is the node's sole remaining input.  \emph{Mass numeral landed (lap 57, judge
pass 17)}: the kernel is re-pinned at \(51/100 \le p_0\) (character-exact to the pass-16
pre-authorization), so \(\varepsilon_0 \ge 1/100 \ge 10^{-4}\) is provable by arithmetic
--- the consumer floor is discharged at the statement level.  \emph{Kernel derived (lap
56)}: whiteness = in-strip \(\wedge\) outside-every-triangle, giving
\(P \ge 1 - P(\text{out of strip}) - P(\text{in a family triangle})\);
\texttt{fpDist\_col\_le} and \texttt{fpDist\_out\_of\_strip\_le} are PROVED (judge-run
clean); the remaining tail is \texttt{fpDist\_any\_triangle\_le} (family-triangle mass
\(\le 1/8\)).
\textbf{ROUTE ESCALATION (lap 57, judge pass 18 CONCURRED):} the formalized
\texttt{F.separated} is \emph{vacuous} at the frozen \(\varepsilon = 10^{-4}\):
\((\tfrac{1}{10}\log 10^4)^2 \approx 0.848 < 1 \le\) any nonzero lattice
distance\(^2\), and node~\ref{X3} proves the clause by exactly that vacuity --- while
the paper's (7.50) whiteness step consumes real separation \(> O(1)\) (``at a distance
\(O(1)\) from \(\Delta\), hence \dots white by Lemma 7.4'', constants independent of
\(\varepsilon\)).  Sound in the paper's \(\varepsilon\)-sufficiently-small regime,
unprovable at our D4 numeral --- \emph{not} a paper error.  Consequences: the pass-17
route ratification of \texttt{fpDist\_any\_triangle\_le} is \textbf{WITHDRAWN}
(adversarial singleton-tiling families are interface-legal; no positive white-mass
bound follows from the interface), and the \texttt{fpDist\_white\_exit\_deep}
statement ratification is \textbf{SUSPENDED} pending the altitude ruling: remedy (A)
shrink \(\varepsilon\) + formalize the real Lemma-7.4 separation, or hybrid
(B+A-small) --- the vertical white gap \texttt{white\_gap\_above\_run\_top}
(\textbf{PROVED} lap 57, judge-run clean: 13 white rows above any run top at the
current \(\varepsilon\), exact-\(\mathbb{Q}\)) plus a smaller shrink for the
horizontal residue.  This node's own statement (\texttt{many\_triangles\_white})
remains ratified; its proof is closed modulo the suspended kernel.  Any D4 change
fires the judge's \(\varepsilon\)-sweep re-ratification tripwire (pass 18 list).
\textbf{ALTITUDE RULING (Trevor, 2026-07-12): Remedy A at \(\varepsilon = 10^{-90}\).}
The D4 numeral shrinks to \(10^{-90}\) --- a rational power of ten (exact-\(\mathbb{Q}\)
machinery preserved; never a transcendental like \(e^{-200}\)) giving separation
\(\tfrac{1}{10}\log 10^{90} = 9\ln 10 \approx 20.7\), comfortably above any plausible
p.48 \(O(1)\).  The kernel route follows p.48 verbatim (paper-faithful ratification);
the real Lemma-7.4 separation becomes content to formalize (node~\ref{X3}'s separation
clause re-opens --- \texttt{sep\_const\_sq\_le\_one} must die); the vertical gap lemma
is kept (holds a fortiori, \(\approx 299\) rows at the new \(\varepsilon\)).  Execution
order: the ManyTriangles split lands first, then a dedicated D4-change lap (numeral +
mechanical repairs only), then separation, then the kernel.  Suspensions lift only via
the post-sweep re-ratification.
\textbf{Pass 23 (2026-07-13): D4 change + real separation EXECUTED} (external Codex
session; split skipped --- still queued): \(\varepsilon = 10^{-90}\) landed with all
mechanical repairs, the full \(\varepsilon\)-sweep re-ratification ran (all seven armed
items discharged; X2/X3/X10 re-verified clean at the new \(\varepsilon\)), and Lemma
7.4's separation is now genuine content (node~\ref{X3}).
\textbf{⚠️ SECOND ALTITUDE-CLASS ESCALATION (pass 23): quantifier order.}  The
foreign-triangle tail \texttt{fpDist\_any\_triangle\_le} was reduced (proved
conditional: \texttt{fpDist\_any\_triangle\_le\_of\_localization\_box}) to the
obligation \(\mathrm{sep} > \sqrt{X^2+Y^2}\) for an explicit X6-localization box ---
but the box proved via the \(16j - 5l\) negative-drift Chernoff has
\(X \approx 2.6\cdot 10^6\), against \(\mathrm{sep} \approx 20.7\).
\textbf{✅ Pass 24 (2026-07-13): escalation DOWNGRADED --- not altitude-class.}  The
p.48 re-read settles it.  The paper's (7.50) \(O(1)\) is a distance \emph{from}
\(\Delta\), not from the start point (the drift slope \(1/4\) is shallower than the
edge slope \(\log 2/\log 9\), so the walk drifts \emph{along} \(\Delta\)), and it is
explicitly \(\varepsilon\)-free --- and the committed geometry already renders exactly
that (top-projection into \(\Delta\) via \texttt{triangle\_top\_mem\_add}, then
\texttt{F.separated}).  The route is sound; the blocker is two throwaway constants.
(i) \texttt{fpDist\_linear\_tail} bounds the \(16j-5l\) MGF with a quadratic penalty
that near-cancels the \(-16\)/step drift, capping the tilt at \(1/20000\) (true ceiling
\(0.213\)) and shipping \(B = 4\cdot 10^7\); the step law's \emph{exact} MGF
(\texttt{geomQuarter} \(\times\) \texttt{pascalNe3}) gives \(B \approx 42\).
(ii) \texttt{fpDist\_height\_tail}'s radius \(Y\) is \emph{existential} (it sums X6's
envelope) --- the real blocker --- but is explicit via renewal-conv \(+\) strictly
increasing heights (\(\Delta l \ge 3\), so each level is visited at most once: renewal
mass \(\le 1\), no renewal theorem) \(+\) \(\Delta l\)'s exact MGF, giving \(Y = 139\).
Box \(= \sqrt{47^2 + 139^2} \approx 147\) vs \(\mathrm{sep} \approx 20.7\): needs one
\emph{cheap} ruling, a numeral re-freeze \(10^{-90} \to 10^{-1000}\)
(\(\mathrm{sep} \approx 230\)).  D4-as-a-parameter is off the table; pass-23's
``\(10^7\)-digit numeral'' was an artifact of the garbage \(B\).
\textbf{🏆 COMPLETE (judge pass 25, 2026-07-13).}  The pass-24 diagnosis executed end to
end in three treadmill laps.  \(B\): \(4\cdot10^7 \to 64\) via an \emph{exact} \texttt{Hold}
MGF closed form (\texttt{tiltZ\_hold\_closed}), freeing the Chernoff tilt from the quadratic
box's \(1/20000\) to \(1/16\) --- and, as predicted, with no \texttt{native\_decide} (the
facts are real-analysis over \(\exp\)).  \(Y\): existential \(\to\) \textbf{150}, via
\texttt{fpDist\_le\_renewal\_conv} \(+\) strictly-increasing heights
(\texttt{renewal\_level\_le\_one}: each height level is visited at most once, so the renewal
mass per level is \(\le 1\) --- no renewal theorem, no local limit law) \(+\) \(\Delta l\)'s
exact MGF; X6's envelope constants were left existential, so the completed node was not
re-opened.  Box \(= \sqrt{51^2+150^2} \approx 158.4\) against \(\mathrm{sep} = 100\ln 10
\approx 230.26\) at the judge's pre-authorized re-freeze \(\varepsilon = 10^{-1000}\) --- it
fits with \(\approx 45\%\) margin.  \texttt{fpDist\_any\_triangle\_le} is therefore a
theorem, \texttt{ManyTriangles.lean} has \emph{zero} sorries, and the pass-18/23
suspension on \texttt{fpDist\_white\_exit\_deep} is lifted in the strongest way available:
it is PROVED.  Dated judge runs 2026-07-13 --- \texttt{fpDist\_any\_triangle\_le},
\texttt{fpDist\_white\_exit\_deep}, \texttt{many\_triangles\_white},
\texttt{fpDist\_linear\_tail\_sharp}, \texttt{renewal\_level\_le\_one} --- all exactly
\([\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}]\).  The second of the
two pinnacle kernels; both are now complete.
\end{lemma}
\begin{proof}
  \leanok
  The depth-gated \(Y\)/\(Z\) induction (\texttt{many\_triangles\_white}, judge pass 16)
  reduces Lemma 7.9 to the single white-exit kernel \texttt{fpDist\_white\_exit\_deep},
  which p.48's (7.50)/(7.51) decomposition splits into the out-of-strip mass
  (\texttt{fpDist\_out\_of\_strip\_le}) and the foreign-triangle mass
  (\texttt{fpDist\_any\_triangle\_le}).  The latter --- blocked since judge pass 18 ---
  follows from \texttt{fpDist\_any\_triangle\_le\_of\_localization\_box} once the X6
  localization box is an explicit numeral and \(\mathrm{sep}\) dominates it: the endpoint
  is projected back onto \(\Delta\) at the top edge (\texttt{triangle\_top\_mem\_add}, via
  the rational inner slope \(5/16 < \log 2/\log 9\), i.e.\ \(9^5 < 2^{16}\)) and Lemma
  7.4's separation (\texttt{F.separated}) excludes every foreign triangle.  The box is
  explicit because (i) the transverse threshold \(B = 64\) comes from the \emph{exact}
  \texttt{Hold} MGF (\texttt{tiltZ\_hold\_closed}) rather than a quadratic bound, and
  (ii) the overshoot radius \(Y = 150\) comes from \texttt{fpDist\_le\_renewal\_conv}
  \(+\) \texttt{renewal\_level\_le\_one} (heights increase by \(\ge 3\), so each level is
  visited at most once) \(+\) \(\Delta l\)'s exact MGF --- leaving X6's envelope constants
  existential.  \(\sqrt{51^2+150^2} \approx 158.4 < 100\ln 10 \approx 230.26\) at
  \(\varepsilon = 10^{-1000}\).
\end{proof}

\begin{lemma}[Large triangles rarely encountered; Lemma 7.10]\label{X10}
  \lean{triangle_encounter_le, fpDistPlus, bigTriangleSet, TriangleFamily.not_mem_two, fpDistPlus_height_tail, fpDistPlus_col_tail, encounter_apex_proximity, encounter_separated_sum}
  \uses{X3, X6, S3}
  \leanok
\textbf{Risk kernel 3b.}  \textbf{Lemma 7.10} (paper (7.60)--(7.65)): after a lengthy crossing,
large triangles are rarely encountered, by a separated-\(\Sigma\) counting argument over the
triangle family (node~\ref{X3}) using the first-passage location law (node~\ref{X6}).  This is
the complementary tail to Case~2.  \emph{Statement pinned + ratified 2026-07-12 (judge pass 8,
paper pp.51--54)}: \texttt{triangle\_encounter\_le} states \(\mathbb{P}(E_{p,s'}) \le
C A^2(1+p)/s' + C e^{-cA^2(1+p)}\) over the endpoint law \(\texttt{fpDistPlus}\;s\;p =
\texttt{fpDist}\;s \star \texttt{iidSum hold}\;p\) --- the D1 encoding of
\(v_{[1,\mathbf{k}+p]}\); the strong-Markov identification is absorbed into the encoding, with
the same trust status as \texttt{fpDist} itself.
\textbf{A-quantifier fix (lap 57, judge pass 18):} the earlier \(\forall A>0\) pin was
\emph{false} --- the (7.61) height display needs \(A^2\) to clear the per-step height
mean \(16\) (\texttt{hold\_mean\_snd}); at fixed small \(A\) and \(p\to\infty\) the
endpoint sits at height \(l_\Delta + \Theta(p)\) outside the \(A^2(1+p)\) window.  The
paper's \(A\) is a standing large parameter.  Old ratification REVOKED; the re-pin
\(\exists A_0 \ge 1,\ \forall A \ge A_0\) (with \(C, c, A_0\) uniform in \(n,\xi,F\))
is \textbf{RATIFIED vs pp.51--52}, consumer-safe (the p.54 union bound instantiates at
one large \(A\); pass-15 read).  The (7.61) tails are pinned separately and ratified as
route pins anchored to the p.52 displays: \texttt{fpDistPlus\_height\_tail}
(\(\le C e^{-cH}\) under margin \(50(1+p) > 16/\text{step}\) drift; the first
\(10(1+p)\) pin was below the drift and self-caught same lap) and
\texttt{fpDistPlus\_col\_tail} (Gaussian width \(\sqrt{1+s}\) + Chernoff under margin
\(10(1+p) > 4/\text{step}\); the \(s^{0.6}\) arithmetic is the assembly site's).  The
(7.65) apex-interval disjointness step (\texttt{TriangleFamily.not\_mem\_two}) is already
\emph{proved} from node~\ref{X3}'s separation --- \emph{disjointness-based, so the
pass-18 route escalation does not touch this node}.
\textbf{Lap 58 (judge pass 19):} both (7.61) tails are \textbf{PROVED} (judge-run clean),
and the assembly is decomposed per pp.52--54: \texttt{encounter\_apex\_proximity} (X10a,
the (7.63)\(\to\)(7.65) confinement --- \textbf{PROVED + ratified vs p.53}, judge-run
clean) + \texttt{encounter\_separated\_sum} (X10b, the p.54 \(\Sigma\)-band sum ---
pinned; the lap-59 regime hypothesis \((s')^2 \le 1+s\) is judge-pre-authorized as
ratified-on-landing: without it the pin is false for \(s' \gg \sqrt{s}\), and the
consumer's standing \(s' \le m^{0.4}\), \(s > m/\log^2 m\) supply it above an absolute
threshold).
\textbf{Lap 59 (judge pass 20): COMPLETE.}  X10b's regime fix landed character-exact to
the pre-authorization (RATIFIED; discharge recorded in the pass-19 addendum) and was
\textbf{PROVED} the same lap; the glue landed in lap 59's final commit, relocating the
pinned headline statement below its engines \emph{character-identically} (judge-diffed
across the move).  Dated judge runs 2026-07-12 on all nine lap-59 declarations ---
X10b, both \(G\)-weight engines, the banded/qualifying steps, the three glue helpers,
and the headline \texttt{triangle\_encounter\_le} itself --- all exactly
\([\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}]\).

\textbf{Judge pass 26 (2026-07-14): a deviation, caught and repaired --- the
\texttt{*\_rpow} engine split.}  An unattended lap (\texttt{61f8e80}) rewrote the deep
hypothesis of \texttt{triangle\_encounter\_le} (and X10a \texttt{encounter\_apex\_proximity})
from \(m/\log^2 m < s\) to \(m^{0.8} < s\), to unblock the Case-3 consumer, whose triangle
sits at row \(n/2-m-1\) (depth \(m+1\)) --- and \(m/\log^2 m < s \Rightarrow
(m+1)/\log^2(m+1) < s\) genuinely \emph{fails} (\(x/\log^2 x\) increasing; fractional-part
counterexample).  The route reasoning was right, but the edit was \emph{not} a generalization:
the two hypotheses are \textbf{incomparable} (\(m^{0.8} \le m/\log^2 m\) only for \(m \gtrsim
10^{15.5}\); below that the new form covers \emph{fewer} \(s\)), and Tao p.51 states Lemma 7.10
with \(s > m/\log^2 m\) \emph{verbatim} --- so the node had stopped rendering the paper's lemma
while staying green and axiom-clean.  Ratifications were REVOKED and the repair mandated:
\textbf{split, don't revert}.  The four weaker-hypothesis lemmas are kept as
\texttt{*\_rpow} \emph{engines} (which is what the Case-3 chain consumes), and the two pins
were \textbf{restored character-identically} and re-proved as corollaries --- \(m \ge 10^{27}\)
via \texttt{log\_sq\_le\_rpow}, \(m < 10^{27}\) trivially (a sub-probability \(\le 1\) against a
bound a large \(C\) makes \(\ge 1\)), exactly the shallow branch this proof already uses.
\textbf{Discharged \texttt{4f51542}}: the judge's character-differ reports \textbf{19/19 pinned
statements byte-identical} to the pre-deviation baseline, and dated runs 2026-07-14 on both
restored pins \emph{and} all four engines are exactly \([\texttt{propext},
\texttt{Classical.choice}, \texttt{Quot.sound}]\).  Net: the campaign gained a strictly stronger
engine \emph{and} kept a faithful Lemma 7.10.  \textbf{Ratifications RESTORED.}
\end{lemma}
\begin{proof}
  \leanok
  Assembly per pp.52--54: trivial branch \(s' < 100C_2A^2(1+p)\) (the \(/s'\) term is
  already \(\ge 1\)); shallow branch \(m < M_{\mathrm{th}} :=
  \max(10^{27}, (S_{0a}+S_{0b}+1)^2)\) absorbed into the \(e^{-cA^2(1+p)}\) term
  (\(M_{\mathrm{th}}\) also clears both abstract thresholds of X10a/X10b); main branch
  splits the encounter indicator pointwise into height-escape + column-escape +
  apex-proximity, discharged by \texttt{fpDistPlus\_height\_tail},
  \texttt{fpDistPlus\_col\_tail}, \texttt{encounter\_apex\_proximity} (X10a) and
  \texttt{encounter\_separated\_sum} (X10b).
\end{proof}

\section{Assembly: Prop 1.17}

\begin{proposition}[Case 3 assembly; Prop 7.8 \(\to\) 7.3 \(\to\) 7.1 \(\to\) Prop 1.17]\label{X11}
  \lean{charFn_decay, key_fourier_decay, Q_black_edge_case3}
  \uses{X9, X10}
  \leanok
\textbf{Case 3} assembly (paper (7.48)/(7.49), (7.52)--(7.67)): the events \(E_*,F_*\), the
choice \(R=\lfloor A^2/\varepsilon^4\rfloor\), and the deterministic claim (7.67), combining
Lemma~7.9 (node~\ref{X9}) and Lemma~7.10 (node~\ref{X10}).  This closes Prop~7.8, hence
Prop~7.3 (node~\ref{X4}) and Prop~7.1, hence the \textbf{crux} \textbf{Proposition 1.17}
(character-sum decay): for every \(A>0\) there is \(C>0\) with, for all \(n\ge 1\) and all
\(\xi\) with \(3\nmid\xi\),
\[
  \Bigl\|\,\mathbb{E}_{Y\sim\texttt{syracZ}\,n}\, e\bigl(-\xi\, Y/3^{n}\bigr)\Bigr\|
    \;\le\; C\,n^{-A}.
\]
This is the single Fourier-analytic input the whole spine reduces to (feeds node~\ref{C10}).
The Case-3 entry statement \texttt{Q\_black\_edge\_case3} ((7.53)--(7.67) interface, budget
hypotheses matching (7.52) with the documented lattice slack) is judge-ratified 2026-07-12
vs pp.48--49 and consumed by the proved \texttt{Q\_black\_edge} split.
\emph{Consumer checks pre-cleared (judge pass 15, pp.48--49 + 54--56 read)}: the (7.54)
0.9m split, the \(R := \lfloor A^2/\varepsilon^4\rfloor\)-after-\(\varepsilon\) choice,
the \(-O(A)\) Markov slack absorbing X9's \(e^{2\varepsilon}\), the depth of every
(7.67) encounter, and X10's consumption at a single large \(A\) at the p.54 union bound
(safe for the \(\exists A_0\) re-pin, judge pass 18).
What remains at ratification time is pinning this node's own statements
(\(E_*, F_*\), (7.66)/(7.67), the assembly chain) when they land.
\lapsrisk{10--20}{high --- assembly volume; consumer geometry pre-cleared}{75\% (2026-07-12 judge pass 15)}
\end{proposition}

\begin{proof}
  \leanok
  PROVED --- \textbf{this closes \S7, and with it the crux Proposition~1.17}.  \emph{Judge
  pass 27 (2026-07-14), dated \texttt{\#print axioms} at pin \texttt{8505bd4}}:
  \texttt{Q\_black\_edge\_case3}, \texttt{charFn\_decay} (Prop~1.17) and
  \texttt{key\_fourier\_decay} (Prop~7.1) all \(=\)
  \([\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}]\).

  The two Case-3 leaves both landed and are themselves axiom-clean:
  \texttt{col\_tail\_mass\_le} ((7.54), the bad-column Gaussian tail) and
  \texttt{few\_white\_mass\_le} ((7.56)).  The latter carries the \emph{paper-faithful} deep
  hypothesis \(m/\log^2 m < s\) --- the weaker \(m^{0.8}\) form introduced during the pass-26
  deviation was confined to the \texttt{\_rpow} \emph{engines} and never reached this leaf.

  \emph{The one step the judge required to see discharged in Lean, not on paper}: the
  Case-3 consumer sits at depth \(m+1\), and \(m/\log^2 m < s \Rightarrow (m+1)/\log^2(m+1) < s\)
  genuinely \textbf{fails}.  The chain closes only by threading a largeness constant, which is
  done in \texttt{Case3.lean:2011--2068}: \((m+1)^{0.8}\le 2m^{0.8}\), \(\log^2 m \le 400\,m^{0.1}\),
  hence \((m+1)^{0.8}\log^2 m \le 800\, m^{0.9}\le m\), so \((m+1)^{0.8}\le m/\log^2 m < s\).
  The largeness is consumed at \(800 \le m^{0.1}\), which needs \(C_{\mathrm{thr}} = 10^{30}\)
  (at \(10^{27}\), \(m^{0.1}\approx 501 < 800\) and the route fails).  Kernel-checked.
\end{proof}
